Hypertension statistics for US adults: an open-source web application for analysis and visualization of US NHANES data

### 2022-11-24 README.Rmd
progressing validation + tutorials
videos and dynamic results

---
output: github_document
---

@@

Our goal is to provide a platform for exploration of the National Health and Nutrition Examination Survey (NHANES) data. We created this website to give context and documentation for users of the application, which can be accessed here: https://bcjaeger.shinyapps.io/nhanesShinyBP/

@@

**Systolic blood pressure by age group in the last three NHANES cycles.**

```{r, echo = FALSE, fig.height=10, fig.width=10}
library(cardioStatsUSA)
nhanes_visualize(outcome_variable = 'bp_sys_mean', 
                 group_variable = 'demo_age_cat',
                 outcome_stats = c("mean"),
                 time_values = c('2013-2014', '2015-2016', '2017-2020'), 
                 title = "",
                 width = 800, 
                 height = 600)[[1]]

@@

**Prevalence of blood pressure control by sex among adults with hypertension.**

```{r, echo = FALSE, fig.height=10, fig.width=10}
nhanes_visualize(outcome_variable = 'bp_control_accaha', 
                 group_variable = 'demo_gender',
                 outcome_stats = c("percentage"),
                 subset_calls = c("htn_accaha" = "Yes"),
                 time_values = 'all', 
                 title = "",
                 width = 800, 
                 height = 600)[[1]]

@@

>>>>

---
output: github_document
always_allow_html: true
---

@@

Our goal is to provide a platform for exploration of the National Health and Nutrition Examination Survey (NHANES) data.

## Web application

- Live version here: https://bcjaeger.shinyapps.io/nhanesShinyBP/

- This website provides context and documentation for our web application

@@

**Figure 1: Systolic blood pressure by age group in the last three NHANES cycles.**

```{r, echo = FALSE, fig.height=10, fig.width=10}
library(cardioStatsUSA)
nhanes_visualize(outcome_variable = 'bp_sys_mean', 
                 group_variable = 'demo_age_cat',
                 outcome_stats = c("mean"),
                 time_values = c('2013-2014', '2015-2016', '2017-2020'), 
                 title = "",
                 width = 800, 
                 height = 600)$fig_object[[1]]

@@

**Figure 2: Prevalence of blood pressure control by sex among adults with hypertension.**

```{r, echo = FALSE, fig.height=10, fig.width=10}
nhanes_visualize(outcome_variable = 'bp_control_accaha', 
                 group_variable = 'demo_gender',
                 outcome_stats = c("percentage"),
                 subset_calls = c("htn_accaha" = "Yes"),
                 time_values = 'all', 
                 title = "",
                 width = 800, 
                 height = 600)$fig_object[[1]]

@@

## Make your own results

You can use the `cardioStatsUSA` web application to run customized analyses of NHANES data. The videos below show how I used it to make Figures 1 and 2.

- This video shows how **Figure 1** was made

```{r, out.width='100%', echo = FALSE}
vembedr::embed_url("https://www.youtube.com/watch?v=SjuzOp4ZMww")
```

- This video shows how **Figure 2** was made

```{r, out.width='100%', echo = FALSE}
vembedr::embed_url("https://www.youtube.com/watch?v=oMporQuiodo")
```

### 2022-11-24 paper/manuscript.Rmd
videos and dynamic results

htn_resistant_2009b <- fig_supp_4_b %>% 
 getElement("fig_data") %>% 
 .[svy_year == '2009-2010' & htn_resistant_accaha == 'Yes'] %>% 
 est_ci()

htn_resistant_2003b <- fig_supp_4_b %>% 
 getElement("fig_data") %>% 
 .[svy_year == '2003-2004' & htn_resistant_accaha == 'Yes'] %>% 
 est_ci()

@@

The age-adjusted prevalence of BP control defined as in the 2017 ACC/AHA BP guideline among non-pregnant US adults with hypertension was lowest in 1999-2000 (`r bracket_drop(bp_control_1999a)`%), highest in 2013-2014 (`r bracket_drop(bp_control_2013a)`%), and `r bracket_drop(bp_control_2017a)`% in 2017-2020 (**Figure 4; Panel A**). Among US adults with hypertension who self-reported taking antihypertensive medication, `r bracket_drop(bp_control_1999b)`%, `r bracket_drop(bp_control_2013b)`%, and `r bracket_drop(bp_control_2017b)`% had controlled BP in 1999-2000, 2013-2014, and 2017-2020, respectively (**Figure 4; Panel B**). The age-adjusted prevalence of resistant hypertension defined as in the 2017 ACC/AHA BP guideline among non-pregnant US adults with hypertension who self-reported taking antihypertensive medication and had ≥1 antihypertensive medication class identified during the medication inventory was lowest in 1999-2000 (`r bracket_drop(htn_resistant_1999a)`%), highest in 2005-2006 (21.4%), and 16.6% in 2017-2020 (**Figure S4; Panel A**). Further restricting this analysis to those who were taking three or more classes of antihypertensive medication, the prevalence of resistant hypertension was lowest in 2009-2010 (64.1%), highest in 2003-2004 (78.3%), and 67.9% in 2017-2020 (**Figure S4; Panel B**).

>>>>

htn_resistant_1999b <- fig_supp_4_b %>% 
 getElement("fig_data") %>% 
 .[svy_year == '1999-2000' & htn_resistant_accaha == 'Yes'] %>% 
 est_ci()

htn_resistant_2005b <- fig_supp_4_b %>% 
 getElement("fig_data") %>% 
 .[svy_year == '2005-2006' & htn_resistant_accaha == 'Yes'] %>% 
 est_ci()

@@

The age-adjusted prevalence of BP control defined as in the 2017 ACC/AHA BP guideline among non-pregnant US adults with hypertension was lowest in 1999-2000 (`r bracket_drop(bp_control_1999a)`%), highest in 2013-2014 (`r bracket_drop(bp_control_2013a)`%), and `r bracket_drop(bp_control_2017a)`% in 2017-2020 (**Figure 4; Panel A**). Among US adults with hypertension who self-reported taking antihypertensive medication, `r bracket_drop(bp_control_1999b)`%, `r bracket_drop(bp_control_2013b)`%, and `r bracket_drop(bp_control_2017b)`% had controlled BP in 1999-2000, 2013-2014, and 2017-2020, respectively (**Figure 4; Panel B**). The age-adjusted prevalence of resistant hypertension defined as in the 2017 ACC/AHA BP guideline among non-pregnant US adults with hypertension who self-reported taking antihypertensive medication and had ≥1 antihypertensive medication class identified during the medication inventory was lowest in 1999-2000 (`r bracket_drop(htn_resistant_1999a)`%), highest in 2005-2006 (`r bracket_drop(htn_resistant_2005a)`%), and `r bracket_drop(htn_resistant_2017a)`% in 2017-2020 (**Figure S4; Panel A**). Further restricting this analysis to those who were taking three or more classes of antihypertensive medication, the prevalence of resistant hypertension was `r bracket_drop(htn_resistant_1999b)`%,  `r bracket_drop(htn_resistant_2005b)`%, and `r bracket_drop(htn_resistant_2017b)`% in 1999-2000, 2005-2006, and 2017-2020, respectively (**Figure S4; Panel B**).
